Drought frequency predicts life history strategies in \emph{Heliophila}

### 2019-03-25 manuscript/manuscript.Rmd
MS updated

Here we combine a long-term global dataset of satellite detected drought events with metadata from natural history collections to test these classic hypotheses within the African endemic mustard genus, _Heliophila_ L. (Brassicaceae). If annuality is an adaptive strategy allowing plants to escape drought prone seasons, then drought frequency should predict the distribution of life history strategies across landscapes, and annual species should be more commonly associated with drought prone regions than perennial species. Additionally, if perenniality offers competative advantage in the absence of drought, associations between life history and drought frequency should be signficant in even if phylogenies are enriched for transitions from annual to perenial life history strategy. Finally, if annual species have adapted to escape predictably drought prone seasons, observations of growing annual species (i.e. occurring in forms other than seed) should be rare during seasons when drought frequency is hightest. Phylogenetic relatedness can influence tests of associations between species’ traits and their environments [@barrett1996comparative;@felsenstein1985phylogenies]. Therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetic context.

@@

$$VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})$$
By combining VCI and TCI, the VHI is distinguishes drought from other forms of vegetative stress [@kogan1995application]. The use of the VHI to detect drought has been validated globally and across ecosystem types [@aghakouchak2015remote], including in southern Africa, the focal region of this study (Figure \@ref:fig(mapsdroughtexamples)). To date, the VHI has most often been applied for evaluating drought risk for agricultural research [e.g., @rojas2011assessing; @kogan2016modelling]. But it also presents a tool to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes [@kerr2003space]. As such, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting intraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought]. Here, we accessed VHI data at $16km^2$ resolution from 1981 to 2015 (https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php) to characterize the seasonal drought frequencies experienced by annual and perennial _Heliophila_ species across their native range of southern Africa.

>>>>

Here we combine a long-term global dataset of satellite detected drought events with metadata from natural history collections to test these classic hypotheses within the African endemic mustard genus, _Heliophila_ L. (Brassicaceae). If annuality is an adaptive strategy allowing plants to escape drought prone seasons, then drought frequency should predict the distribution of life history strategies across landscapes, and annual species should be more commonly associated with drought prone regions than perennial species. Additionally, if perenniality offers competative advantage in the absence of drought, associations between life history and drought frequency should be signficant in even if phylogenies are enriched for transitions from annual to perenial life history strategy. Finally, if annual species have adapted to escape predictably drought prone seasons, observations of growing annual species (i.e. occurring in forms other than seed) should be rare during seasons when drought frequency is highest. Phylogenetic relatedness can influence tests of associations between species’ traits and their environments [@barrett1996comparative;@felsenstein1985phylogenies]. Therefore we assessed the relationship between life history distribution and drought frequency in a phylogenetic context.

@@

$$VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})$$
By combining VCI and TCI, the VHI is distinguishes drought from other forms of vegetative stress [@kogan1995application]. The use of the VHI to detect drought has been validated globally and across ecosystem types [@aghakouchak2015remote], including in southern Africa, the focal region of this study (Figure \@ref(fig:mapsdroughtexamples)). To date, the VHI has most often been applied for evaluating drought risk for agricultural research [e.g., @rojas2011assessing; @kogan2016modelling]. But it also presents a tool to study seasonal patterns in the frequency of drought across environments and to test hypotheses about the effect of drought on ecological and evolutionary processes [@kerr2003space]. As such, the VHI  has been applied recently to study drought related ecology of natural species and proven useful for predicting intraspecific variation in drought tolerance traits and genes [@Mojica2016;@dittberner2018natural;@monroe2018drought]. Here, we accessed VHI data at $16km^2$ resolution from 1981 to 2015 (https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php) to characterize the seasonal drought frequencies experienced by annual and perennial _Heliophila_ species across their native range of southern Africa.

### 2019-03-25 manuscript/manuscript.tex
MS updated

Here we combine a long-term global dataset of satellite detected drought
events with metadata from natural history collections to test these
classic hypotheses within the African endemic mustard genus,
\emph{Heliophila} L. (Brassicaceae). If annuality is an adaptive
strategy allowing plants to escape drought prone seasons, then drought
frequency should predict the distribution of life history strategies
across landscapes, and annual species should be more commonly associated
with drought prone regions than perennial species. Additionally, if
perenniality offers competative advantage in the absence of drought,
associations between life history and drought frequency should be
signficant in even if phylogenies are enriched for transitions from
annual to perenial life history strategy. Finally, if annual species
have adapted to escape predictably drought prone seasons, observations
of growing annual species (i.e.~occurring in forms other than seed)
should be rare during seasons when drought frequency is hightest.
Phylogenetic relatedness can influence tests of associations between
species' traits and their environments (Felsenstein, 1985; Barrett
\emph{et al.}, 1996). Therefore we assessed the relationship between
life history distribution and drought frequency in a phylogenetic
context.

@@

\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\] By combining VCI
and TCI, the VHI is distinguishes drought from other forms of vegetative
stress (Kogan, 1995b). The use of the VHI to detect drought has been
validated globally and across ecosystem types (AghaKouchak \emph{et
al.}, 2015), including in southern Africa, the focal region of this
study (Figure @ref:fig(mapsdroughtexamples)). To date, the VHI has most
often been applied for evaluating drought risk for agricultural research
(e.g., Rojas \emph{et al.}, 2011; Kogan \emph{et al.}, 2016). But it
also presents a tool to study seasonal patterns in the frequency of
drought across environments and to test hypotheses about the effect of
drought on ecological and evolutionary processes (Kerr \& Ostrovsky,
2003). As such, the VHI has been applied recently to study drought
related ecology of natural species and proven useful for predicting
intraspecific variation in drought tolerance traits and genes (Mojica
\emph{et al.}, 2016; Dittberner \emph{et al.}, 2018; Monroe \emph{et
al.}, 2018b). Here, we accessed VHI data at \(16km^2\) resolution from
1981 to 2015
(\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
characterize the seasonal drought frequencies experienced by annual and
perennial \emph{Heliophila} species across their native range of
southern Africa.

>>>>

Here we combine a long-term global dataset of satellite detected drought
events with metadata from natural history collections to test these
classic hypotheses within the African endemic mustard genus,
\emph{Heliophila} L. (Brassicaceae). If annuality is an adaptive
strategy allowing plants to escape drought prone seasons, then drought
frequency should predict the distribution of life history strategies
across landscapes, and annual species should be more commonly associated
with drought prone regions than perennial species. Additionally, if
perenniality offers competative advantage in the absence of drought,
associations between life history and drought frequency should be
signficant in even if phylogenies are enriched for transitions from
annual to perenial life history strategy. Finally, if annual species
have adapted to escape predictably drought prone seasons, observations
of growing annual species (i.e.~occurring in forms other than seed)
should be rare during seasons when drought frequency is highest.
Phylogenetic relatedness can influence tests of associations between
species' traits and their environments (Felsenstein, 1985; Barrett
\emph{et al.}, 1996). Therefore we assessed the relationship between
life history distribution and drought frequency in a phylogenetic
context.

@@

\[VHI_{y,w,i} = 0.5(VCI_{y,w,i}) + 0.5(TCI_{y,w,i})\] By combining VCI
and TCI, the VHI is distinguishes drought from other forms of vegetative
stress (Kogan, 1995b). The use of the VHI to detect drought has been
validated globally and across ecosystem types (AghaKouchak \emph{et
al.}, 2015), including in southern Africa, the focal region of this
study (Figure \ref{fig:mapsdroughtexamples}). To date, the VHI has most
often been applied for evaluating drought risk for agricultural research
(e.g., Rojas \emph{et al.}, 2011; Kogan \emph{et al.}, 2016). But it
also presents a tool to study seasonal patterns in the frequency of
drought across environments and to test hypotheses about the effect of
drought on ecological and evolutionary processes (Kerr \& Ostrovsky,
2003). As such, the VHI has been applied recently to study drought
related ecology of natural species and proven useful for predicting
intraspecific variation in drought tolerance traits and genes (Mojica
\emph{et al.}, 2016; Dittberner \emph{et al.}, 2018; Monroe \emph{et
al.}, 2018b). Here, we accessed VHI data at \(16km^2\) resolution from
1981 to 2015
(\url{https://www.star.nesdis.noaa.gov/smcd/emb/vci/VH/vh_ftp.php}) to
characterize the seasonal drought frequencies experienced by annual and
perennial \emph{Heliophila} species across their native range of
southern Africa.
